---
title: "Debugging with error handling"
---

When writing scripts, you rarely want a future run with new data to crash and emit an error message like:

::: {.cell-output .cell-output-stderr}
Warning message:
NAs introduced by coercion 
:::

The code ran fine before! Did you run the script with a typo in a parameter? Or is there badly formatted data in your new input file? Which of the 10,000 rows has the bad data??

If you're developing a tool or planning to reuse your script, you can save a lot of headaches in the long-run by writing your code to validate inputs and fail early with bespoke, informative messages.

For instance, say you have a function that simply reads a csv and writes a new one with rounded values from a single column:

```{r}
round_column <- function(df, col_num_raw, prec_raw) {
  col_num <- as.numeric(col_num_raw)
  prec <- as.numeric(prec_raw)
  df$rounded <- round(as.numeric(df[[col_num]]), digits = prec)
  print(df)
}
```

You might run this function with a test dataset:

```{r}
test_data_1 <- data.frame(
  labels = c("A", "B", "C"),
  values = c(10.556, 21.891, 3.14159)
)
print(test_data_1)
results_1 <- round_column(test_data_1, 2, 2)
```

And celebrate that it works!

```{r}
print(results_1)
```

But then you try some real data:

```{r}
test_data_2 <- data.frame(
  labels = c("A", "B", "C"),
  values = c(10.556, 21.891, '3.141S9')
)
print(test_data_2)
results_2 <- round_column(test_data_2, 2, 2)
```

Well that's vague!