---
title: "Merge, Join, and Bind"
---

We have two data frames, `df1` and `df2`, that we want to combine.
```{r}
df1 <- data.frame(id = 1:3, letter = c("A", "B", "C"))
print(df1)
df2 <- data.frame(id = 2:4, value = c(10, 20, 30))
print(df2)
```

In base R, we use `merge()`:

```{r}
merged_df <- merge(df1, df2, by = "id")
print(merged_df)
```

The default behavior of `merge()` is an inner join, which means that only rows with matching `id` values in both data frames are included in the result. To get the same result using the tidyverse, we can use `dplyr::inner_join()`:

```{r}
library(dplyr)
inner_join_df <- inner_join(df1, df2, by = "id")
print(inner_join_df)
```

If we want to include all rows from `df1` and only matching rows from `df2`, we can use the `all.x = TRUE` argument in `merge()`:

```{r}
left_join_df <- merge(df1, df2, by = "id", all.x = TRUE)
print(left_join_df)
```

And in the tidyverse, we can use `dplyr::left_join()`:
```{r}
left_join_df_tidy <- left_join(df1, df2, by = "id")
print(left_join_df_tidy)
```

If we want to include all rows from `df2` and only matching rows from `df1`, we can use the `all.y = TRUE` argument in `merge()`:

```{r}
right_join_df <- merge(df1, df2, by = "id", all.y = TRUE)
print(right_join_df)
```

And in the tidyverse, we can use `dplyr::right_join()`:
```{r}
right_join_df_tidy <- right_join(df1, df2, by = "id")
print(right_join_df_tidy)
```

If we want to include all rows from both `df1` and `df2`, we can use the `all = TRUE` argument in `merge()`:

```{r}
full_join_df <- merge(df1, df2, by = "id", all = TRUE)
print(full_join_df)
```

And in the tidyverse, we can use `dplyr::full_join()`:
```{r}
full_join_df_tidy <- full_join(df1, df2, by = "id")
print(full_join_df_tidy)
```

Merge automatically sorts the result by the `by` column(s). If you want to keep the original order, you can use the `sort = FALSE` argument in `merge()`. The dplyr functions do not automatically sort the result. 

When both data.frames have columns with the same name, both `merge()` and the dplyr join functions will add suffixes to the column names to distinguish them. You can specify your own suffixes using the `suffixes` argument in `merge()` or the `suffix` argument in the dplyr join functions.

When the columns used for merging have different names in the two data frames, you can use the `by.x` and `by.y` arguments in `merge()` or the `by` argument in the dplyr join functions to specify which columns to use for merging.

`merge()` can be slow for large datasets, so some people recommend dplyr for performance reasons. However, I've seen benchmarks showing that base R is considerably faster. To get real speed improvements, though, you can also use the `data.table` package. It is far faster than either base R or dplyr, and has a `merge()` function with syntax similar to base R's `merge()`. `data.table` also has other highly-optimized functions, like `fread()`, which are indispensable for working with large datasets.

#TODO: do some benchmarks right in this document

